Portfolios of Subgraph Isomorphism Algorithms

### 2015-12-06 paper/paper.tex

\begin{abstract}
Subgraph isomorphism is a computationally challenging problem with lots of important practical applications, for example in computer vision, biochemistry, and model checking. There are a number of state-of-the-art algorithms for solving it, each of which has its own performance characteristics. As with many other hard problems, the single best choice of algorithm overall is rarely the best algorithm for each instance. We develop a portfolio approach that leverages novel features to characterise subgraph isomorphism problems to dynamically decide which algorithm to use on a per-instance basis. We demonstrate significant performance improvements on a large set of hard benchmark problems. In addition, we show how algorithm selection models can be leveraged to gain new insights into what affects the performance of an algorithm.\end{abstract}

@@

In this paper, we report experimental results on a larger set of 8 algorithms, including new variants, and a larger benchmark. We show that 5 of these algorithms are single best algorithms, depending on the CPU time limit considered, and that combining preprocessings with an algorithm selection on an instance by instance basis allows us to achieve better overall performance than any single algorithm.

@@

We consider the following exact algorithms:
 VF2 \cite{Cordella:2004}, which is a simple backtracking search algorithm, especially fast on trivially satisfiable instances;
 LAD \cite{Solnon:2010}, which  does very strong reasoning based on local all different redundant constraints;
 and the Glasgow algorithm \cite{McCreesh:2015}, which does an expensive preprocessing pass once, followed by weaker reasoning during search.
 The Glasgow algorithm has a parameter, which controls the lengths of paths used when reasoning about non-adjacent vertices.  In experiments reported in \cite{McCreesh:2015}, the choice of paths of length 3 was used as a reasonable compromise---longer paths lead to prohibitively expensive preprocessing on larger, denser instances. This is
often not the best choice on an instance by instance basis: sometimes path-based reasoning gives no benefit at all, sometimes considering only paths of length 2 suffices, occasionally paths of length 4 are helpful, and even looking at paths of length 3 is relatively expensive on some graphs. Therefore, in this new study, we consider 4 different instances of the Glasgow algorithm, with path lengths varying from 1 to 4.

@@

We consider 5725 instances which are available in a simple text format\footnote{\url{http://liris.cnrs.fr/csolnon/SIP.html}}. These instances are grouped into 12 classes.
\begin{itemize}
\item Class 1 contains randomly generated scale-free graphs \cite{}.
\item Classes 2 and 3 contain various kinds of graphs coming from \cite{}: class 2 contains small instances generated from the 50 first graphs of the benchmark, whereas class 3 contains larger instances where pattern graphs are coming from class 2, and target graphs are coming from the next 50 graphs of the benchmark.
\item Classes 4 to 8 contain randomly generated graphs coming from \cite{GraphDatabase}: bounded-degree graphs for classes 3 and 4, regular meshes for classes 5 and 6, and random graphs with uniform edge probabilities for class 7.
\item Classes 9 and 10 contain instances coming from images \cite{}.
\item Class 11 contains instances coming from meshes modeling 3D objects \cite{}.
\item Class 12 contains random graph instances belonging to the phase transition.
\end{itemize}
Note that Classes 3 and 12 were not considered in the previous experimental study reported in \cite{McCreesh:2015}.

>>>>

\begin{abstract}
Subgraph isomorphism is a computationally challenging problem with lots of important practical
applications, for example in computer vision, biochemistry, and model checking. There are a number
of state-of-the-art algorithms for solving it, each of which has its own performance
characteristics. As with many other hard problems, the single best choice of algorithm overall is
rarely the best algorithm for each instance. We develop a portfolio approach that leverages novel
features to characterise subgraph isomorphism problems to dynamically decide which algorithm to use
on a per-instance basis. We demonstrate significant performance improvements on a large set of hard
benchmark problems. In addition, we show how algorithm selection models can be leveraged to gain new
insights into what affects the performance of an algorithm.
\end{abstract}

@@

In this paper, we report experimental results on a larger set of 8 algorithms, including new
variants, and a larger benchmark. We show that 5 of these algorithms are single best algorithms,
depending on the CPU time limit considered, and that combining preprocessing with an algorithm
selection on an instance by instance basis allows us to achieve better overall performance than any
single algorithm.

\section{Background}

\todo[inline]{Some background on subgraph isomorphism algorithms, maybe move
some stuff from the intro here.}

The per-instance algorithm selection problem~\cite{rice_algorithm_1976} is to select from an
algorithm portfolio~\cite{huberman_economics_1997,gomes_algorithm_2001} the one expected to perform
best on a given problem instance. Algorithm selection systems usually build machine learning models
of the algorithms or the portfolio they are contained in to forecast which algorithm to use in a
particular context. Using the model predictions, one or more algorithms from the portfolio are
selected to be run sequentially or in parallel.

Here, we consider the case where exactly one algorithm is selected for solving the problem. One of
the most prominent and successful systems that employs this approach is
SATzilla~\cite{xu_satzilla_2008}, which defined the state of the art in SAT solving for a number of
years. Other application areas include constraint solving~\cite{omahony_using_2008}, the travelling
salesperson problem~\cite{kotthoff_improving_2015}, and AI planning~\cite{seipp_learning_2012}.

The interested reader is referred to a recent survey~\cite{kotthoff_algorithm_2014} for additional
information on algorithm selection.

@@

We consider the following exact algorithms: VF2 \cite{Cordella:2004}, which is a simple backtracking
search algorithm, especially fast on trivially satisfiable instances; LAD \cite{Solnon:2010}, which
does very strong reasoning based on local all different redundant constraints; and the Glasgow
algorithm \cite{McCreesh:2015}, which does an expensive preprocessing pass once, followed by weaker
reasoning during search.

 The Glasgow algorithm has a parameter, which controls the lengths of paths used when reasoning
 about non-adjacent vertices.  In experiments reported in \cite{McCreesh:2015}, the choice of paths
 of length 3 was used as a reasonable compromise---longer paths lead to prohibitively expensive
 preprocessing on larger, denser instances. This is often not the best choice on an instance by
 instance basis: sometimes path-based reasoning gives no benefit at all, sometimes considering only
 paths of length 2 suffices, occasionally paths of length 4 are helpful, and even looking at paths
 of length 3 is relatively expensive on some graphs. Therefore, in this new study, we consider 4
 different instances of the Glasgow algorithm, with path lengths varying from 1 to 4.

@@

We consider 5725 instances which are available in a simple text
format\footnote{\url{http://liris.cnrs.fr/csolnon/SIP.html}}. These instances are grouped into 12
classes.

\begin{itemize}
\item Class 1 contains randomly generated scale-free graphs \cite{}.
\item Classes 2 and 3 contain various kinds of graphs coming from \cite{}: class 2 contains small instances generated from the 50 first graphs of the benchmark, whereas class 3 contains larger instances where pattern graphs are coming from class 2, and target graphs are coming from the next 50 graphs of the benchmark.
\item Classes 4 to 8 contain randomly generated graphs coming from \cite{GraphDatabase}: bounded-degree graphs for classes 3 and 4, regular meshes for classes 5 and 6, and random graphs with uniform edge probabilities for class 7.
\item Classes 9 and 10 contain instances coming from images \cite{}.
\item Class 11 contains instances coming from meshes modeling 3D objects \cite{}.
\item Class 12 contains random graph instances belonging to the phase transition.
\end{itemize}

\noindent Randomly generated instances with particular properties:

\begin{itemize}
    \item Bounded-degree random graphs (family 3, 4) \cite{GraphDatabase}
    \item Random 4d mesh graphs (family 5, 6) \cite{GraphDatabase}
    \item Random graphs (family 7) \cite{GraphDatabase}
    \item Randomly generated scale-free graphs. (family 1) \cite{Solnon:2010}
    \item Phase transition (family 12)
\end{itemize}

We consider a total of 5725 instances.

\subsubsection{Features}

\todo[inline]{Describe features here.}

This gives us a total of 30 features.

The data we use in this paper is available as Aslib scenario GRAPHS-2015.

\subsection{Algorithm Selection approach}

We apply two static steps to each instance before solving it. First, the
computation of the LAD features can detect unsatisfiability of an instance. If
an instance is identified as unsatisfiable, we do not process it further.
Second, we run the VF2 solver for 0.05 seconds. This solves easy instances
without the overhead of running algorithm selection and avoids potentially
making incorrect solver choices.

Computing the LAD features detects unsatisfiability on 1919 instances. Running
VF2 for 0.05 seconds removes a further 1470. We are left with 2336 hard
instances that we consider for algorithm selection.

To decide which solver to run on the remaining instances, we use
LLAMA~\cite{kotthoff_llama_2013}. In preliminary experiments, we identified
random forest regression to predict the performance difference between pairs of
algorithms, with the best algorithm identified as the one with the best
cumulative performance difference over all pairs, as the best approach. We make
no claims that this is the overall best approach.

\section{Results}

Table~\ref{tab:res} shows the performance of our algorithm selection approach,
compared to two baselines. The virtual best solver is the oracle predictor that,
for each instance, chooses the best solver from our portfolio. This is the upper
bound of what an algorithm selection approach can achieve. The single best
solver is the one solver from the portfolio that has the overall best
performance across the entire set of instances. We consider it a lower bound on
performance.

\begin{table}[ht]
\centering
\begin{tabular}{lr}
  \toprule
& mean runtime\\
  \midrule
virtual best & 5822.809\\
  single best & 7782.569\\
  LLAMA & 6595.275\\
   \bottomrule
\end{tabular}
\vspace{1ex}
\caption{Algorithm selection performance on reduced set of 2336
instances.}\label{tab:res}
\end{table}

We are able to close more than 60\% of the performance gap between the single
best and the virtual best solver, our lower and upper bounds. For comparison, we
show the results achieved on the entire set of instances without
unsatisfiability detection and presolving in Table~\ref{tab:resfull}. We achieve
similar results with more than 58\% of the gap between the single best and the
virtual best closed.

\begin{table}[ht]
\centering
\begin{tabular}{lr}
  \toprule
& mean runtime\\
  \midrule
virtual best & 2375.913\\
  single best & 3177.569\\
  LLAMA & 2706.330\\
   \bottomrule
\end{tabular}
\vspace{1ex}
\caption{Algorithm selection performance on the full set of 5725
instances.}\label{tab:resfull}
\end{table}

Figure~\ref{fig:scatter} shows detailed results for each of the 2336 instances
in the reduced set. Our algorithm selection approach is by no means able to
identify the best solver on a majority of instances, although we are able to
achieve performance improvements overall. There seems to be no particular
pattern as to where algorithm selection performs well and where it fails.

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{figures/perfScatter}
\caption{Single best algorithm vs.\ portfolio performance on all instances by
family. Points above the diagonal indicate that the portfolio achieved better
performance, while the single best solver was better on instances below the
diagonal.}
\label{fig:scatter}
\end{figure}

Note that Classes 3 and 12 were not considered in the previous experimental study reported in \cite{McCreesh:2015}.
